Staged Abstract Interpreters

### 2019-02-26 paper/20_bg.tex
paper
migrate to scalaz

In this section, we first describe the abstract syntax of the language we will
analyze, then present the generic interface shared between the four different
interpreters. We first instantiate the interpreter to the concrete one; then we
proceed in two directions: by staging it we obtain a staged concrete interpreter
(Section~\ref{stagedinterp}), by abstracting it we obtain an unstaged abstract
interpreter (Section~\ref{unstaged_abs});

\input{21_bg_interp}
\input{23_bg_stage}
\input{22_bg_abs}

>>>>

In this section, we first describe the abstract syntax of the language we will
analyze, then present the generic interface shared between the four different
interpreters. We choose to use Scala to demonstrate the idea, but the idea of
utilizing monads and multi-stages also can be easily constructed in other functional
languages, such as MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
  DBLP:conf/flops/Kiselyov14} or Template Haskell \cite{Sheard:2002:TMH:636517.636528}.
We first instantiate the interpreter to the concrete one; then we
proceed in two directions: by staging it we obtain a staged concrete interpreter
(Section~\ref{stagedinterp}); by abstracting it we obtain an unstaged abstract
interpreter (Section~\ref{unstaged_abs}).

\input{21_bg_interp}
\input{22_bg_stage}
\input{23_bg_abs}

### 2019-02-26 paper/21_bg_interp.tex
paper

We consider a call-by-value $\lambda$-calculus in direct-style, but extended
with numbers, arithmetics, recursions and conditionals. Other effectful features
such as assignments also can be supported easily; in Section~\ref{cases_imp}, we
will add more imperative features to the language. Since we are mostly
interested in analyzing the dynamic behaviors of the program, we disguised any
static semantics and type system. We also assume that input programs are
well-typed and all variables are distinct. The abstract syntax is shown as
follows:

@@

We give the concrete semantics of \TLang by showing a big-step definitional
interpreter. The interpreter is a recursive function that takes the program AST,
environment and store, and returns the evaluated value and the accompanied
store. The environment is a mapping from identifiers to addresses, and the store
is a mapping from addresses to values. We use the store to model recursion and
mutation in concrete semantics; it is also useful for polyvariant analysis. This
environment-and-store-passing style big-step interpreter is standard and can
also be obtained by refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03,
Wei:2018:RAA:3243631.3236800} a small-step CESK machine
\cite{DBLP:conf/popl/FelleisenF87}.

>>>>

We consider a call-by-value $\lambda$-calculus in direct-style, extended
with numbers, arithmetics, recursions and conditionals. Other effectful features
such as assignments can also be supported readily. In Section~\ref{cases_imp}, we
will add more imperative features to the language. Since we are mostly
interested in analyzing the dynamic behaviors of the program, we disguised any
static semantics and type system. We also assume that input programs are
well-typed and all variables are distinct. The abstract syntax is shown as
follows:

@@

We will give the concrete semantics using a big-step definitional
interpreter. The interpreter is a recursive function that takes the program AST,
environment and store, and returns the evaluated value and the accompanied
store. The environment is a mapping from identifiers to addresses, and the store
is a mapping from addresses to values. We use the store to model recursion and
mutation in concrete semantics; it is also useful for polyvariant analysis. This
environment-and-store-passing style big-step interpreter is standard and can
also be obtained by refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03,
Wei:2018:RAA:3243631.3236800} a small-step CESK machine
\cite{DBLP:conf/popl/FelleisenF87}.

### 2019-02-26 paper/22_bg_stage.tex
paper

Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
multi-stage programming framework implemented as a Scala library that enables
dynamic code generation in a type-safe manner. Different from the syntactic
approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations,
LMS distinguishes binding-time solely based on types. LMS provides a type
constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an
expression will be known at next stage. All operations acting on a @Rep[T]@
expression will be residualized in the generated code. However, we need to
provide intermediate representation and code generation support for type @T@ to
LMS. Fortunately, the LMS framework already provides such support for primitive
types and commonly used data structure such as arrays and maps; implementing
such support for custom class is also straightforward. Let's go back and see how
LMS can be used to specialize the power function we mentioned in Section 1.

>>>>

Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
multi-stage programming framework implemented as a Scala library that enables
dynamic code generation in a type-safe manner. Different from the syntactic
approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14,
DBLP:conf/gpce/CalcagnoTHL03} that uses quotations, LMS distinguishes
binding-time solely based on types. LMS provides a type constructor @Rep[T]@
where @T@ can be an arbitrary type, indicating an expression will be known at
next stage. All operations acting on a @Rep[T]@ expression will be residualized
in the generated code. However, we need to provide intermediate representation
and code generation support for type @T@ to LMS. Fortunately, the LMS framework
already provides such support for primitive types and commonly used data
structure such as arrays and maps; implementing such support for custom class is
also straightforward. Let's go back and see how LMS can be used to specialize
the power function we mentioned in Section 1.
